Holomorphic field theories and higher algebra

### 2023-08-16 highervertexops.tex
merge

\def\g{{\mathfrak g}}
\def\U{{\rm U}}
\def\mc{\mathcal}
\def\mcol{\, | \,}
\def\ot{\otimes}
\def\Disj{\operatorname{Disj}}
\def\Open{\operatorname{Open}}
\def\Vect{\operatorname{Vect}}
\def\jou{{\mathtt{j}}}
\def\bgamma{{\mathbf{\gamma}}}
\def\bbeta{{\mathbf{\beta}}}
\def\CDO{{\cC\cD}}
\def\SU{{\rm SU}}
\def\sfa{\mathsf{a}}
\def\del{\partial}
\def\til{\Tilde}
\def\Sol{\text{Sol}}
\def\Spec{\text{Spec}}
\def\C{{\mathbb{C}}}
\def\RHom{{\rm RHom}}
%\def\sfR{\mathsf{R}}
%\def\sfA{\mathsf{A}}

@@

Consider now the space of all polynomials
\[
\C[z_i,\lambda_i] = \Sym(V_z \oplus V_\lambda) \cong \Sym(V_z) \otimes \Sym(V_\lambda).
\]
The Lie algebra $\sl(2)$ acts by the lowering and raising operators
\beqn
f = z_2 \del_{z_1} - \lambda_1 \del_{\lambda_2} \quad\text{and}\quad e = z_1 \del_{z_2} - \lambda_2 \del_{\lambda_1}.
\eeqn
The space of polynomials breaks up into a direct sum over the subrepresentations 
\[
V_{a,b} = \Sym^a(V_z) \otimes \Sym^b(V_\lambda)
\]
of polynomials that are homogenous of order $a$ in the variables $z_i$ and homogenous of order $b$ in the variables~$\lambda_i$.

@@

Consider now  
\[
|(z,\lambda)|^2 \define z_1 \lambda_1 + z_2 \lambda_2, 
\]
which describes the evaluation pairing. 
It is preserved by the $SL(2)$-action, and hence we see that the canonical quotient map 
\[
q: \C[z_i,\lambda_i] \to R = \C[z_i,\lambda_i]/(|(z,\lambda)|^2 = 1)
\]
is $SL(2)$-equivariant.
Moreover, inside $V_{a,b}$, any multiple of $|(z,\lambda)|^2$ is annihilated by $q$,
so $R$ is spanned by the images of $H_{a,b}$.
Thus, there is a direct sum decomposition $\sfR = \bigoplus_{a,b} \sfR_{a,b}$ where $\sfR_{a,b} = q(H_{a,b})$ is the image of the subspace of harmonic polynomials in $\C[z_i,\lambda_i]$ under the quotient map.

@@

\begin{lem}
Each $\sfR_{a,b}$ is isomorphic to the irreducible $SL(2)$-representation of dimension $(a+b+1)$.
Its highest weight vector is $z_1^a \lambda_2^b \in \sfR_{a,b}$, and a basis for $\sfR_{a,b}$ is obtained as
\beqn
\{z_1^a \lambda_2^b, f (z_1^a \lambda_2^b), f^2 (z_1^a \lambda_2^b), \ldots, f^{a+b} (z_1^a \lambda_2^b) \} .
\eeqn
\end{lem}

@@

>>>>

\def\g{{\mathfrak g}}
\def\U{{\rm U}}
\def\mc{\mathcal}
\def\mcol{\, | \,}
\def\ot{\otimes}
\def\Disj{\operatorname{Disj}}
\def\Open{\operatorname{Open}}
\def\Vect{\operatorname{Vect}}
\def\jou{{\mathtt{j}}}
\def\bgamma{{\mathbf{\gamma}}}
\def\bbeta{{\mathbf{\beta}}}
\def\CDO{{\cC\cD}}
\def\SU{{\rm SU}}
\def\sfa{\mathsf{a}}
\def\del{\partial}
\def\til{\Tilde}
\def\Sol{\text{Sol}}
\def\Spec{\text{Spec}}
\def\C{{\mathbb{C}}}
\def\RHom{{\rm RHom}}
\newcommand{\sff}{\mathsf{f}}
\newcommand{\sfe}{\mathsf{e}}
%\def\sfR{\mathsf{R}}
%\def\sfA{\mathsf{A}}

@@

Consider now the space of all polynomials
\[
\C[z_i,\lambda_i] = \Sym(V_z \oplus V_\lambda) \cong \Sym(V_z) \otimes \Sym(V_\lambda).
\]
The Lie algebra $\sl(2)$ acts by the lowering and raising operators
\beqn
\sff = z_2 \del_{z_1} - \lambda_1 \del_{\lambda_2} \quad\text{and}\quad \sfe = z_1 \del_{z_2} - \lambda_2 \del_{\lambda_1}.
\eeqn
The space of polynomials breaks up into a direct sum over the subrepresentations 
\[
V_{a,b} = \Sym^a(V_z) \otimes \Sym^b(V_\lambda)
\]
of polynomials that are homogenous of order $a$ in the variables $z_i$ and homogenous of order $b$ in the variables~$\lambda_i$.

@@

Consider now  
\[
|(z,\lambda)|^2 \define z_1 \lambda_1 + z_2 \lambda_2, 
\]
which describes the evaluation pairing. 
It is preserved by the $SL(2)$-action, and hence we see that the canonical quotient map 
\[
q \colon \C[z_i,\lambda_i] \to R = \C[z_i,\lambda_i]/(|(z,\lambda)|^2 = 1)
\]
is $SL(2)$-equivariant.
Moreover, inside $V_{a,b}$, any multiple of $|(z,\lambda)|^2$ is annihilated by $q$,
so $R$ is spanned by the images of $H_{a,b}$.
Thus, there is a direct sum decomposition $\sfR = \bigoplus_{a,b} \sfR_{a,b}$ where $\sfR_{a,b} = q(H_{a,b})$ is the image of the subspace of harmonic polynomials in $\C[z_i,\lambda_i]$ under the quotient map.

@@

\begin{lem}
Each $\sfR_{a,b}$ is isomorphic to the irreducible $SL(2)$-representation of dimension $(a+b+1)$.
Its highest weight vector is $z_1^a \lambda_2^b \in \sfR_{a,b}$, and a basis for $\sfR_{a,b}$ is obtained as
\beqn
\{z_1^a \lambda_2^b, \sff (z_1^a \lambda_2^b), \sff^2 (z_1^a \lambda_2^b), \ldots, \sff^{a+b} (z_1^a \lambda_2^b) \} .
\eeqn
\end{lem}

@@

\subsection{Residues}

Define
\beqn
\op{Res} \colon \sfA \to \C[1]
\eeqn
by the rules that $\op{Res}|_{\sfA^0} = 0$ and
\beqn
\op{Res} \left(f(z,\lambda) \omega \right) = p (f(z,\lambda)) (z=0) .
\eeqn

\begin{lem} The following are true regarding the higher algebraic residue:
\begin{enumerate}
\item The residue is $\lie{sl}(2)$ invariant.
\item For $f \in \C[z_1,z_2]$ one has $\op{Res}(f \omega) = f(0)$.
\item For any $f \in \sfA$ one has $\op{Res}(\dbar f) = 0$.
\end{enumerate}
\end{lem}
\begin{proof}
Item (1) holds since $p$ is the $\lie{sl}(2)$-equivariant projection. 
For (2), note that for $f \in \C[z_1,z_2]$ that $p(f) = f$.

Finally, observe that for $f = z_1^a \lambda_2^b$ we have
\beqn
\op{Res}(\dbar f) = \op{Res}(b z_1^{a+1} \lambda_2^{b-1}) .
\eeqn
For $b =0$ this expression clearly vanishes. 
For $b = 1$ the right hand side is $p( z_1^{a+1})(0) = 0$ for all $a \geq 0$.
Finally, for $b > 1$ one has $p(z_1^{a+1} \lambda_2^{b-1}) = 0$ for all $a \geq 0$.
Using $\lie{sl}(2)$ invariance this proves (3).
\end{proof}
